---
title: "Salmon Group 15"
author: "Anastasia Kunz, Zoe Duerksen-Salm, Catherine Takata"
date: "1/30/2021"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: hide 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(feasts)
library(tsibble)
library(slider)
library(lubridate)
library(patchwork)
```

# Overview:
This “Overview” section (above the tabs) should contain, in whatever order you choose: 
An engaging image (with caption, including photo credit) that is relevant to the dataset
A brief summary (3 - 4 sentences) of the dataset, and what is included in this “report”
A map of the fish ladder location (you can make this in R on your own, or include an existing map appropriately licensed, with attribution)
A professionally formatted data citation
Remember: All code that you used to wrangle the data and prepare the graphs should be available to see if we click on the Code buttons. 

```{r}
#initial data read in

fish_data <- read_csv(here("data","willamette_fish_passage.csv")) %>% 
  clean_names()
```
# Tab 1: Time Series


```{r timeseries}

#data wrangling: convert date, create tsibble, select species, na to 0
ts_salmon <- fish_data %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  as_tsibble(key = NULL, index = date) %>% 
  select(date, coho, jack_coho, steelhead) %>% 
  mutate_at(c(2:4), ~replace(., is.na(.), 0)) 

#Steelhead Plot

steelhead_plot <-ggplot(ts_salmon, aes(x = date, y = steelhead))+
  geom_line(color = "indianred")+
  labs(
    y = "Steelhead",
    x = " "
  )+
  theme_minimal()+
  ylim(0, 1250)

# Coho Plot
coho_plot <- ggplot(ts_salmon, aes(x = date, y = coho))+
  geom_line(color = "rosybrown")+
  theme_minimal()+
  labs(
    y = "Coho",
    x = "Date"
  )+
  ylim(0, 1250)

# Jack Coho Plot
jackcoho_plot <- ggplot(ts_salmon, aes(x = date, y = jack_coho))+
  geom_line(color = "lightsalmon1")+
  theme_minimal()+
  labs(
    y = "Jack Coho",
    x = " "
  )+
  ylim(0, 1250)


```

```{r, fig.cap = "**Figure 1.** Shows the number of adult Steelhead (top plot), Jack Coho (middle plot), and Coho (bottom plot) salmon passing through the Willamette Falls Fish Ladder on the Willamette River in Oregon from January 1 2001 - December 31 2010."}
# Final graph using patchwork

salmon_patchwork <-(steelhead_plot / jackcoho_plot / coho_plot)

salmon_patchwork + plot_annotation(
  title = 'Salmon Passage by Species',
  subtitle = 'Number of Steelhead, Coho, and Jack Coho Salmon counted at the Willamette Falls Fish Ladder \n2001 - 2010'
)
```

### Summary
By tracking the number of adult salmon passing through the fish ladder, species specific trends become clear:

- Steelhead, Jack Coho, and Coho all have song seasonal passage patterns, with Coho and Jack Coho passing in the fall, while Steelhead pass over a greater span of time in the summer months.

- Adult Steelhead usually pass in higher numbers than Coho and Jack Coho, except in 2009 and 2010 when Coho numbers surpassed both Steelhead and Jack Coho.

- Coho show an increasing trend with increasing numbers of adults passing in recent years.

- Jack Coho are showing relatively stable low numbers, with no strong increasing trend.


# Tab 2: Seasonplots

This “Seasonplots” tab should contain: 

- A finalized seasonplot for each species (coho, jack coho, steelhead). You can either facet, or use patchwork, to make the figure. Add a caption. 
- Add 2 - 3 bulletpoints summarizing the major trends you see in the seasonplots.
Remember: All code that you used to wrangle the data and prepare the graphs should be available to see if we click on the Code buttons. 
```{r seasonplot1}
coho_jack_steel <- fish_data %>% 
  select(date, coho, jack_coho, steelhead) 

# convert df to a tsibble
cjs_ts <- coho_jack_steel %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  as_tsibble(key = NULL, index = date)
```

```{r seasonplot2}
cjs_ts_tidy <- cjs_ts %>% 
  pivot_longer(`coho`:`steelhead`, # The columns I'm gathering together
               names_to = "fish_species", # new column name for existing names
               values_to = "fish_count") # new column name to store values

cjs_ts_tidy %>% 
  filter(year(date) > 2000) %>% 
  gg_season(y = fish_count) +
  theme_minimal() +
  labs(x = "Time (month)",
       y = "Total monthly fish count (n)",
       title = "Seasonal Fish Counts")
```
